---
title: "Exploratory Data Analysis"
author: "Amy Gill"
date: "January 20, 2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

  - methods/analysis section that explains the process and techniques used, such as data cleaning, data exploration and visualization, any insights gained, and your modeling approach


# Exploratory Data Analysis and Modeling Methods

This document performs exploratory data analysis (EDA) on the EdX MovieLens dataset `Rdata/edx.Rdata` generated in `data-download.Rmd`.

## Data import

```{r}
load("Rdata/edX.Rdata")
```

## Exploratory Data Analysis

### Overview

```{r}
dim(edx)
```

The `edx` dataset contains 9,000,055 observations of each variable.

```{r}
summary(edx)
```

The variables consist of `userId`, `movieId`, `rating`, `timestamp`, `title`, and `genres`. There are no missing values in the dataset.

### Rating: The `rating` variable

As the goal of this study is to predict rating based on other features, I first inspect the `rating` variable.
```{r}
typeof(edx$rating)
```

```{r}
factor(edx$rating) %>% levels()
```

The rating variable conists of a numeric five-star rating with half-star increments. All ratings in the dataset conform to this pattern. As some operations will work best with numeric values but other will require categorical values, I add the factor-coded rating as a separate variable in `edx` named `ratingFactor`.

```{r}
edx <- edx %>%
  mutate(ratingFactor = factor(rating))

head(edx)
```

```{r}
summary(edx$rating)
```

The mean rating is 3.512 and the median rating is 4. The IQR is 1, or 1 star.

```{r}
edx %>%
  group_by(rating) %>%
  summarize(count = n(), probability = count/nrow(edx))
```

The two most frequent ratings are 4 stars (28.7%) and 3 stars (23.6%). The two least frequent ratings are 0.5 stars (0.9%) and 1.5 stars (1.1%).


```{r}
edx %>%
  ggplot(aes(x = ratingFactor)) +
  geom_bar()
```

There is a clear discretization effect where whole-star ratings are more frequent than half-star ratings. The distribution appears to be skewed to the right.

```{r}
edx %>%
  filter(ratingFactor %in% c(1, 2, 3, 4, 5)) %>%
  ggplot(aes(x = ratingFactor)) +
  geom_bar()
```
```{r}
edx %>%
  filter(ratingFactor %in% c(0.5, 1.5, 2.5, 3.5, 4.5)) %>%
  ggplot(aes(x = ratingFactor)) +
  geom_bar()
```

When considering whole-star and half-star ratings independently, they show similar trends suggesting a distribution centered between 3.5 and 4. However, the discretization to whole-star values is distinct and will need to be included in a highly accurate model.

### Users: The `userId` Variable

`userId` encodes a unique integer ID number for each user. As these integers serve only as identifiers and have no numerical meaning, I convert them to factors:

```{r}
edx <- edx %>%
  mutate(userId = factor(userId))

length(levels(edx$userId))
```

There are 69878 unique users in the dataset. The `users` data frame will contain useful reference metrics for each user.

```{r}
users <- edx %>%
  group_by(userId) %>%
  summarize(reviews = n(),
            avgRating = mean(rating),
            sdRating = sd(rating),
            medianRating = median(rating),
            prob_0.5 = mean(rating == 0.5),
            prob_1 = mean(rating == 1),
            prob_1.5 = mean(rating == 1.5),
            prob_2 = mean(rating == 2),
            prob_2.5 = mean(rating == 2.5),
            prob_3 = mean(rating == 3),
            prob_3.5 = mean(rating == 3.5),
            prob_4 = mean(rating == 4),
            prob_4.5 = mean(rating == 4.5),
            prob_5 = mean(rating == 5))

head(users, 10)
```

